Analyzing the Impact of Social Attributes on Commit Integration Success

### 2017-03-25 Paper/MSRChallenge2017.tex
changed incorrect sentence

In particular, we analyzed contributions in the 
context of continuous integration, using the technical factors and one social
factor found in the challenge
corpus~\cite{msr17challenge}, and the social factors found on a developer's
GitHub profile.  We used the technical factors to estimate the complexity of the
commit, and the social factors to estimate the developer's experience.
Overall, our study investigates how well different technical and social factors predict
contribution quality. This information can help core team members prioritize project
contributions, and will also provide a further understand of quality enforcement
in the software development process.

We find that social factors are an important aspect of build success but
these factors alone do not fully predict if a contribution will be successfully 
integrated into the project. We find
that core team members are more likely to achieve build success than non-core
members, and that developers with more followers are more likely to submit commits that pass the
continuous integration build.

@@

\vspace{1ex}
\noindent\textbf{Goal.}
The goal of this study is to identify and understand the relationship 
between the social factors that describe the
developer's experience (stars, followers, etc.) and technical properties that
describe the commit's complexity
(files added, files modified, etc.)
when predicting the likelihood a developer's commit
to be considered successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.

\vspace{1ex}
\noindent\textbf{Question.}
%Previous studies have identified a relationship between technical and social 
%factors and pull request acceptance. 
We are interested in understanding whether
social factors, used as a proxy for experience, can be used to predict
a different success measure: commit integration success, therefore giving
developers and integrators another tool to be able to identify the quality
of a contribution performed by a particular developer.
%
Our main research question, upon which we can build by investigating particular attributes, is:

\textbf{Do social attributes, gathered 
from the GitHub profiles of developers,
allow for more accurate predictions of 
the developer's build success
than only using technical data?}

\vspace{1ex}
\noindent\textbf{Metric.} We measure quality software contributions based on the build status of the
commit. We used a decision tree classifier~\cite{Quinlan86} to identify
important factors in predicting build success;  decision tress are well
established  and interpretable. In particular, we use a J48 decision tree because this classification mechanism has the ability to identify the attributes that discriminate the various instances most clearly, and we are interested in analyzing what particular features can predict with higher accuracy build success.
We first build a
decision tree classifier using only technical factors, which highlights
the degree to which commit complexity predicts build integration
success.  We then added the social factors to identify the degree to which they
add predictive power to a the initial technical model. 
We use 10-fold cross validation across our evaluation.

@@

\vspace{1ex}
\noindent\textbf{Technical Factors} We collected technical factors from the 
Mining Software Repositories (MSR) Challenge data set~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy a commit's
complexity, such as commit size. 
The technical factors used are:
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn, changed lines of code.
\item Files added by this commit.
\item Files modified by this commit.
\item Tests added by this commit.
\item Tests deleted by this commit.
\item Total source files in project at the time of the commit.
\item Source lines of source code in the project at
	the time of the commit.
\item Travis status, or status of the Travis build integration.
\end{itemize}

@@

\begin{itemize}
\item Number of repositories the developer can access, including those
the developer owns, contributes
to, and for which the developer has organizational membership access.%
\footnote{https://developer.github.com/v3/repos/}
\item  Number of projects the developer has stared. Starring a project
to shows approval of the repository and creates  a bookmark for later 
access.\footnote{https://help.github.com/articles/about-stars/}
\item Number of developers that follow the author. 
\item Number of developers that the author follows .
\item Number of contributions made by the author in the last year.
\end{itemize}

These social factors are all factors that are tied to a GitHub user's profile
and that are generally correlated with user time and engagement with the site.
Thus, we use them as a proxy for
experience. If we were unable to find information about an author than we
assigned the commit empty values for these factors.
We also added one item from the MSR Challenge data set, indicating
whether the commit's developer is a core member  
of the associated project. These values were investigated to understand and be able to quantify the developers’ experience.

@@

Our data set~\cite{msr17challenge} consists
of commits that date as far back as 2011. The social data we 
gather is recent, gathered in January 2017. To mitigate the threat of
using future results to predict previous events,
we removed only include data set commits from the years 2014--2016.  Our intuition is
that the social attributes are less likely to have substantively changed more recently.

Since we are interested in the success of the job status of each contribution, we 
converted all Travis build statuses to two possible values:
\emph{successful} (passed in the original provided data set) and \emph{unsuccessful}
(errored, failed, canceled
and started in the original data set). We also condensed down all passing and
failing builds in a commit to a single pass or fail for each build status.  If
the commit contained both passing and failing build statuses, we treat it as a failed build.

74.43\% of the builds in the reduced dataset were successful, and thus a naive
classifier that always guesses ``successful'' would succeed 74.43\% of the time.
To penalize misclassification for
both classes equally, we created a second balanced dataset by randomly 
sub-sampling the successful builds until the two build status were
equal, a process known as downsampling. We downsampled the entire dataset being analyzed and then performed cross-validation on it. Since the cross-validation is stratified, the proportion of 50-50 is maintained through the training and testing folds. We train and present results for both
data sets individually in Sections \ref{eval}A, \ref{eval}B and \ref{eval}C.
\section{Evaluation} \label{eval}
\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Non-Downsampled}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
	\textbf{Correctly Classified Instances}&  285882 (77.01\%) & 301068
	(81.10\%)  & 121694 (64.10\%) & 144441 (76.09\%)\\

@@

\hline
\end{tabular}
\\
\center
	\caption{\label{resultsTable}Comparison of predictive power by running 10-fold
    cross validation of two decision trees with a non-downsampled dataset (2nd and 3rd column) and downsampled (4th and 5th column). 
	%The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}
	%Zack: the above sentence isn't true for the second row.
	}   
\end{table*}
To address our main research question: ``Do social attributes, gathered 
from the GitHub profiles of developers,
allow for more accurate predictions than only using technical data?", we 
created a decision tree classifier~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using 10-fold cross 
validation with the data in each data set.

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.  Without downsampling, the 
decision tree shows an increase of 4.09\% \textit{correctly classified} instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
 After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

@@

Based on the decision tree results, we noticed that core team members where
important when classifying build success, so we decided to investigate if core
team members have greater build success rates.
We classified the build success rate for
core team members and non-core team members using the data set without 
downsampling.  We found that core team members
triggered 89.39\% (331827/371208) of the builds in the data set and 75.58\%
(250808/331827) of those builds
passed.  Non-core team members triggered 10.61\% (39381/371208) of the builds 
in the data set
and 64.70\% (25479/39381) of the builds passed.  This shows that core team 
members are more
likely to have a successful build integration than non-core team members. We hypothesize the reason for this to be that core team members are more familiar with the code, therefore are more likely to have their builds pass.

@@

We also investigated whether developers with followers were more likely to have a
successful build integration, since followers may be seen as vouching for
the developer's abilities.  We specifically investigated if a developer with 
followers was more likely to pass the build in the data set without 
downsampling.  We first investigated whether developers with one or more followers
were more likely to pass the build than developers without any followers, to
verify the assumption that followers vouch for a developer's abilities. To
investigate this question, we removed the commits without author profile
information (20374 builds which reduced the total build count from 371208 to
350834). We found 
that only 0.87\% (3059/350834) of the
contributions in the data set were performed by developers with zero 
followers (excluding the developers in which we were not able to extract profile
information for). We found that 99.13\% (347775/350834) of the contributions 
were performed by
developers with at least one or more followers.
Developers with 0 followers had a successful build commit 68.49\% (2095/3059)
of the time.  Developers with one or more followers were successful 74.78\%
(260056/347775) of the time.

We also investigated developers with a large number of followers (1000 or more
followers).  We found that 6.94\% (24347/350834) of the contributions in the 
data set were
performed by developers with 1000 or more followers.  We found that these
contributions had a 78.61\% (19138/24347) success rate.  These results show 
developers
with more followers are more likely to have a successful integration build. Reasons for this might be

@@

Overall, we have found that social attributes which proxy experience, such if
the developer is a core contributor for the project, produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors
as the ones studied in this paper are significant predictors when
evaluating the predictive power of build commit success.

The build success by core team member results show that core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commit are not thoroughly
scrutinized, although further studied will be required to draw definite
conclusions.

The follower results indicate that developers with more followers are more
likely to pass the build, although followers are not a perfect indicator.
Developers with over 1000 followers had more than a 10\% increase compared to
developers with no followers.  This shows that higher quality developers are more
likely to have a large number of followers.

@@

 The scripts we have created to adapt and extract the datasets we needed, and the 
 scripts to obtain the developer social data are publicly available to facilitate 
 the transparency and quality of the process\footnote{https://goo.gl/CqTX44}.

@@

\section{Related work}
The most similar related work to this paper is a case study investigating how
social, technical, coordination, and prior-certification factors affect build
success for a project with 40 developers at IBM Toronto Labs~\cite{Hassan06}.  This work 
focused on one project and measured experience as the number of contributions to
that project while our study focuses on a wide range of open source projects and
measures experience through developers' profile information.
Another study investigated the relationship between authors, files in project,
and team organization when predicting build
success on an IBM project with 151 developers.  The study found that as the
number of authors or files in a build increases, the probability of build
success decreases~\cite{Kwan11}. A separate investigation used a dataset from the
same project and
found that team structures affected the teams' build success rate~\cite{Wolf09}.
Similar studies have shown the relationship between both social and technical attributes attributes and 
different measures of success: pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse}, pull request evaluation latency~\cite{Yu15}.
Tsay et al.~\cite{tsay14icse} 
show that the social connection between the developer initiating a
pull request and the project integrators is a key factor in predicting 
acceptance of a pull 
request. Unlike this previous study, in our study, we look at a broader aspect of 
social aspects, and focus on predicting the success in Travis
CI jobs. Gousios et al.~\cite{gousios14} found that the change location of a commit was
an important indicator in the acceptance rate of those commits by the project.
Also, the inclusion of test cases with a commit has been shown to increase the
commit's acceptance rate~\cite{gousios15}.
Bettenburgh et 
al.~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, Vasilescu et al.~\cite{vasilescu14} explore how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated into the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in GitHub and its correlation with software 
quality. By contrast with their work, we focus on the relationship between independent social 
attributes and continuous integration success.

>>>>

In particular, we analyzed contributions in the 
context of continuous integration, using technical factors and one social
factor from the challenge
corpus~\cite{msr17challenge}, and social factors from a developer's
GitHub profile.  We used the technical factors to estimate the commit's
complexity, and the social factors to estimate the developer's experience.
Overall, our study investigates how different technical and social factors predict
contribution quality. This information can help core team members prioritize project
contributions, and will also provide a further understand of quality enforcement
in the software development process.

We find that social factors are an important aspect of build success but
these factors alone do not fully predict whether a contribution will be successfully 
pass the build tests. We find
that core team members are more likely to achieve build success than non-core
members, and that developers with more followers are more likely to submit commits that pass the
continuous integration build.

@@

\vspace{1ex}
\noindent\textbf{Goal.}
The goal of this study is to identify and understand the relationship 
between the social factors that describe the
developer's experience (stars, followers, etc.) and technical properties that
describe the commit's complexity
(files added, files modified, etc.)
when predicting the likelihood that a developer's commit
will pass the continuous integration build tests for the project.

\vspace{1ex}
\noindent\textbf{Question.}
%Previous studies have identified a relationship between technical and social 
%factors and pull request acceptance. 
We are interested in understanding whether
social factors, used as a proxy for experience, can be used to predict
commit integration success. If so, 
developers and integrators will have another tool to estimate the quality
of a developer's contribution.
%
Our main research question is:

\textbf{Do social attributes, gathered 
from developers' GitHub profiles ,
allow for more accurate predictions of 
the developer's build success
than only technical attributes?}

If social attributes improve the prediction of build success, then we are
interested in investigating how certain attributes, such as the developer's
follower count, correlates with build success.

\vspace{1ex}
\noindent\textbf{Metric.} We measure quality software contributions based on the
build status of the
commit. We used a decision tree classifier~\cite{Quinlan86} to identify
important factors in predicting build success;  decision tress are well
established and interpretable. In particular, we use a J48 decision tree because
this classifier has the ability to identify which attributes 
discriminate the various instances most clearly, and we are interested in
analyzing which features can be used to predict build success.
We first build a
decision tree classifier using only technical factors to determine
how well commit complexity predicts build integration
success.  We then include the social factors in the classification problem to 
identify the degree to which the social factors improve the model's predictive
power.
We evaluate the models with 10-fold cross-validation.

@@

\vspace{1ex}
\noindent\textbf{Technical Factors.} We collected technical factors from the 
Mining Software Repositories (MSR) Challenge data set~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy a commit's
complexity, such as commit size. 
The technical factors used are:
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn, changed lines of code.
\item Files added by this commit.
\item Files modified by this commit.
\item Tests added by this commit.
\item Tests deleted by this commit.
\item Total source files in project at the time of the commit.
\item Source lines of source code in the project at
	the time of the commit.
\item Travis status, or status of the Travis build integration.
\end{itemize}

@@

\begin{itemize}
	\item Number of repositories the developer can access (including those
the developer owns, contributes
		to, and for which the developer has organizational membership
		access).%
\footnote{https://developer.github.com/v3/repos/}
\item  Number of projects the developer has stared. Starring a project
to shows approval of the repository and creates a bookmark for later 
access.\footnote{https://help.github.com/articles/about-stars/}
\item Number of developers that follow the author. 
\item Number of developers that the author follows.
\item Number of contributions made by the author in the last year.
\end{itemize}

These social factors are all factors that are tied to a GitHub user's profile
and that are generally correlated with user time and engagement with the site.
Thus, we use them as a proxy for
experience. If we were unable to find information about an author, then we
assigned the commit empty values for these factors.
We also added one item from the MSR Challenge data set, indicating
whether the commit's developer is a core member  
of the associated project. These values were investigated to estimate developers’ experience.

@@

The challenge dataset~\cite{msr17challenge} consists
of commits that date as far back as 2011. The social data we 
gathered is accurate for January 2017. To mitigate the threat of
using future results to predict previous events,
we removed only include data set commits from the years 2014--2016.  Our intuition is
that these social attributes are unlikely to have significant recent changes.

Since we are interested in the success of the job status of each contribution, we 
converted all Travis build statuses to two possible values:
\emph{successful} (passed in the challenge dataset~\cite{msr17challenge}) and \emph{unsuccessful}
(errored, failed, canceled
and started in the original data set). We also condensed all passing and
failing builds in a commit to a single pass or fail for each build status.  If
the commit contained both passing and failing build statuses, we treat it as
both a passing and failed build.

74.43\% of the builds in the reduced dataset were successful, and thus a naive
classifier that always guesses ``successful'' would succeed 74.43\% of the time.
To penalize misclassification for
both classes equally, we created a second balanced dataset by randomly 
sub-sampling the successful builds until the two build status were
equal, through a process known as downsampling. We downsampled the entire
dataset and then trained decision tree classifiers.  We validated the
classifiers with cross-validation. Since the cross-validation is stratified, the
proportion of 50-50 is maintained through the training and testing folds for the
downsampled data. We present results for both
datasets in Section \ref{eval}A,  and only the full dataset for Sections \ref{eval}B and \ref{eval}C.

\section{Evaluation} \label{eval}
\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Non-Downsampled}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
	\textbf{Correctly Classified Instances}&  285882 (77.01\%) & 301068
	(81.10\%)  & 121694 (64.10\%) & 144441 (76.09\%)\\

@@

\hline
\end{tabular}
\\
\center
	\caption{\label{resultsTable}Comparison of predictive power by running 10-fold
    cross-validation of two decision trees with a non-downsampled dataset
	(2\textsuperscript{nd} and 3\textsuperscript{rd} column) and downsampled
	(4\textsuperscript{th} and 5\textsuperscript{th} column). 
	%The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}
	%Zack: the above sentence isn't true for the second row.
	}   
\end{table*}
To address our main research question: ``Do social attributes, gathered 
from the GitHub profiles of developers,
allow for more accurate predictions than only using technical data?", we 
created a decision tree classifier~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using 10-fold 
cross-validation with the data in each data set.

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.  Without downsampling, the 
decision tree shows an increase of 4.09\% \textit{correctly classified} instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
 After including the social data
in the dataset, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

@@

Based on the decision tree results, we noticed that core team members were
important when classifying build success, so we decided to investigate if core
team members have greater build success rates.
We classified the build success rate for
core team members and non-core team members using the data set without 
downsampling.  We found that core team members
triggered 89.39\% (331827/371208) of the builds in the data set and 75.58\%
(250808/331827) of those builds
passed.  Non-core team members triggered 10.61\% (39381/371208) of the builds 
in the data set
and 64.70\% (25479/39381) of the builds passed.  This shows that core team 
members are more
likely to have a successful build integration than non-core team members. We hypothesize the reason for this to be that core team members are more familiar with the code, therefore are more likely to have their builds pass.

@@

We also investigated whether developers with followers were more likely to have a
successful build integration, since followers may be seen as vouching for
the developer's abilities.  We specifically investigated if a developer with 
followers was more likely to pass the build in the data set without 
downsampling.  We first investigated whether developers with one or more followers
were more likely to pass the build than developers without any followers, to
verify the assumption that followers vouch for a developer's abilities. To
investigate this question, we removed the commits without author profile
information (20374 builds which reduced the total build count from 371208 to
350834). We found 
that only 0.87\% (3059/350834) of the
contributions in the data set were performed by developers with zero 
followers. We found that 99.13\% (347775/350834) of the contributions 
were performed by
developers with at least one or more followers.
Developers with 0 followers had a successful build commit 68.49\% (2095/3059)
of the time.  Developers with one or more followers were successful 74.78\%
(260056/347775) of the time.

We also investigated developers with a large number of followers (1000 or more
followers).  We found that 6.94\% (24347/350834) of the contributions in the 
data set were
performed by developers with 1000 or more followers.  We found that these
contributions had a 78.61\% (19138/24347) success rate.  These results show 
developers
with more followers are more likely to have a successful integration build.
These
result may be due to the fact that developers with followers are more likely to
have open source development experience, and the developers' experience
influenced the build success rate..

@@

Overall, we have found that social attributes which proxy experience, such as if
the developer is a core project contributor , produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors,
such 
as the ones studied in this paper, are significant predictors when
evaluating the predictive power of build commit success.

The build success by core team member results show that core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commits are not as
scrutinized as core team member commits before the integration build tests, although further studied will be required to draw definite
conclusions.

The follower results indicate that developers with more followers are more
likely to pass the build, although followers are not a perfect indicator.
Developers with over 1000 followers had more than a 10\% increase compared to
developers with no followers.  This shows that higher quality open source developers are more
likely to have a large number of followers.

@@

 The scripts we have created to create the datasets and the 
 scripts to obtain the developer social data are publicly available 
 \footnote{https://goo.gl/CqTX44}.

@@

\section{Related work}
The most similar related work to this paper is a case study investigating how
social, technical, coordination, and prior-certification factors affect build
success for a project with 40 developers at IBM Toronto Labs~\cite{Hassan06}.  This work 
focused on one project and measured experience as the number of contributions to
the project while our study focuses on a wide range of open source projects and
measures experience through developers' profile information.
Another study investigated the relationship between authors, files in project,
and team organization when predicting build
success on an IBM project with 151 developers.  The study found that as the
number of authors or files in a build increases, the probability of build
success decreases~\cite{Kwan11}. A separate investigation used a dataset from the
same project and
found that team structures affected the teams' build success rate~\cite{Wolf09}.
Similar studies have shown the relationship between both social and technical attributes attributes and 
different measures of success: pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse}, pull request evaluation latency~\cite{Yu15}.
Tsay et al.~\cite{tsay14icse} 
show that the social connection between the developer initiating a
pull request and the project integrators is a key factor in predicting 
acceptance of a pull 
request. Unlike this previous study, in our study, we look at a broader aspect of 
social aspects, and focus on predicting the success in Travis
CI jobs. Gousios et al.~\cite{gousios14} found that the change location of a commit was
an important indicator in the acceptance rate of those commits by the project.
Also, the inclusion of test cases with a commit has been shown to increase the
commit's acceptance rate~\cite{gousios15}.
Bettenburgh et 
al.~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, Vasilescu et al.~\cite{vasilescu14} explore how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated into the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in GitHub and its correlation with software 
quality. By contrast with their work, we focus on the relationship between independent social 
attributes and continuous integration success.
